---
title: "Introduction to projrsimple"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{intro}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(projrsimple)
```

## Introduction

The goal of `projrsimple` is to make it easy to run a clean project workflow.

It helps you initialise a reproducible project structure with pre‐defined
directories, run analysis scripts, and ensure clean output directories
with minimal effort.

## Installation

You can install the development version of `projrsimple` like so:

``` r
if (!requireNamespace("remotes", quietly = TRUE)) {
  install.packages("remotes")
}
remotes::install_github("MiguelRodo/projrsimple")
```

## TL;DR

1. Install `projrsimple` (see above).
2. Open an R session in your project directory (folder where you want to work).
3. Run `projr_init()` to initialise your project.
4. Write code in scripts (`.R`, `.Rmd`, `.qmd` and/or Quarto projects) in the project directory.
5. Run `projr_run` to execute all scripts. Save outputs to `_output` directory.
6. View rendered documents in the `docs` directory.

To connect to GitHub automatically, run `projr_init(init_github = TRUE)` (step 3) *after* the following steps:

1. Create a GitHub account ([link](https://www.github.com)).
2. Set up a personal access token (PAT) in R ([instructions](https://happygitwithr.com/https-pat#tldr)).

## Details

### Project Initialisation with `projr_init`

The function `projr_init` is designed to help you quickly set up a standard project  
structure.
It will (all of these features are optional):

- **Create Standard Directories:**  

  By default, it creates directories for:
  - Raw data: `_raw_data`
  - Rendered documents: `docs`
  - Final outputs: `_output`
  - Temporary/cache files: `_tmp`
  - Raw (non‐rendered) docs: `_reference`
  
  You can supply your own values (provide `NULL` to skip a directory).  

- **Initialise a Git Repository:**  

  `projr_init` can initialise a Git repository in your project.
  It will automatically add the  cache directory (by default `_tmp`) to your
  `.gitignore` to avoid committing  
  temporary files, and (if enabled) will add and commit any changes.  
  *Note:* To push your repository to GitHub, you’ll need a GitHub account and must  
  ensure that R is configured with your personal access token (PAT).

- **Create a README File:**  

  A default README is created with instructions on how to run your analyses
  using `projrsimple`, as well as a brief description of the project,
  its structure and empty fields for links and project details.

- **Connect to GitHub:**

  `projr_init` can connect your local Git repository to GitHub.
  The repository is private by default.

### Running Analyses with `projr_run`

After you have initialised your project, you can use `projr_run` to execute your  
analyses. By default:

- **Script Detection and Execution:**  

  All scripts in your working directory with extensions `.R`, `.Rmd`/`.rmd`, and  
  `.qmd` are run automatically. If a Quarto project is detected
  (`_quarto.yml` exists), then a Quarto project is rendered.

- **Copying Generated Documents:**  

  The function will copy any generated documents (typically `.html`, `.pdf`, or  
  `.docx`) to the `docs` directory.

- **Output Directory Management:**  

  Optionally, the output directory (by default `_output`) is cleared before running  
  all the scripts.
  By default, this is `FALSE` (to avoid unintended deletions), but should
  be set to `TRUE` to ensure clean runs.

- **Selective Execution:**  

  You can specify a subset of scripts to run, rather
  than all found in the `scripts` directory.

### Multiple Pipelines

For long-running and/or complex workflows, you can run `projr_run` multiple times to create  
separate "pipelines".
For example, you might run one pipeline that processes raw data  
and another that generates reports, each using different script selections and  
output/document directories.
This modular approach allows you to flexibly manage  
complex projects with multiple stages.

For example:

```{r, eval=FALSE}
# processed data
projr_run(
  scripts = "process-data.R",
  dir_output = "_output/processed_data",
  dir_docs = "docs/processed_data",
  clear_output_and_docs = TRUE
)
# analyse data
projr_run(
  scripts = "analyse-data.R",
  dir_output = "_output/analysis",
  dir_docs = "docs/analysis",
  clear_output_and_docs = TRUE
)
```

Since separate `output` and `docs` directories are used for the different pipelines,
the analysis pipeline can be run cleanly (clear `output` and `docs` directories) without
affecting the processed data pipeline.

### Execution Directory

Keeping scripts in a sub-directory (e.g. `src/` or `scripts/`)
can improve organisation, but may complicate the execution directory
when running `Quarto` and `RMarkdown` documents.

The execution directory is the directory where `R` thinks it is when running scripts.

For `Quarto` and `RMarkdown`, the execution directory is the directory of the `.Rmd`/`.qmd` file.
So, if we run `quarto::quarto_render("scripts/analysis.qmd")` from the project directory,
the execution directory is actually `scripts/`.
For example, if we run `quarto::quarto_render("scripts/analysis.qmd")` from the project directory,
the execution directory is `scripts/`.

This has two possible problems:

- The user needs to be aware of this, and specify all paths
within the `Quarto`/`RMarkdown` doc relative to `scripts/`,
rather than the project directory. This is not typically what people
expect.
- This is not the same behaviour as `R` scripts. If we run `source("scripts/analysis.R")`,
the execution directory is not `scripts/` but the project directory.

For this reason, to prevent surprises, if you use
`Quarto` or `RMarkdown` it is recommended to keep all scripts in the
project directory.

#### Workarounds

However, if you would like to keep `Quarto`/`RMarkdown` docs in a subdirectory,
you have the following two workarounds:

##### Option 1 (preferred): Use `knitr` and specify the execution directory

To ensure consistency between `projr_run` and manual knitting, 
place the following at the very top of your `.Rmd`/`.qmd` file
(within a setup chunk, of course):

```{r, eval=FALSE}
knitr::opts_knit$set(root.dir = "/your/desired/path")
```

For example, if the `Quarto` doc is in the `scripts` directory but
you want the execution directory to be the project directory, replace
`"/your/desired/path"` with `".."` (note two dots - this means "one level up").

##### Option 2: Specify the execution directory in `projr_run`

You can specify the `dir_exec` argument in `projr_run` to set the execution directory.
For example:

```{r, eval=FALSE}
projr_run(
  scripts = c("scripts/data_processing.R", "scripts/analysis.qmd"),
  dir_exec = "." # note only one dot
)
```

This forces the execution directory to be the project directory for 
both the `R` script and `Quarto` doc, even though the
scripts are in a subdirectory.

The problem with this approach is that the `Knit`/`Render` button
in `RStudio` will still render the `Quarto`/`RMarkdown` docs
with the parent directory as the execution directory (`scripts/`, 
in this case).
This is the reason `Option 1` is preferred.

On the other hand, if you wanted the `R` script `scripts/data_processing.R`
to have `scripts/` as the execution directory, then using `projr_run` 
with `dir_exec = "scripts/"` would be a good approach.

## Citation

To cite `projrsimple` in publications, use:

> **Miguel Rodo** (2024). *projrsimple: Initialise and run a simple project workflow.* Version 1.0.0. Available at: [https://github.com/MiguelRodo/projrsimple](https://github.com/MiguelRodo/projrsimple/#readme).

Alternatively, in **BibTeX** format:

```bibtex
@Misc{rodo,
  title = {projrsimple: Initialise and run a simple project workflow},
  author = {Miguel Rodo},
  url = {https://github.com/MiguelRodo/projrsimple/#readme},
  abstract = {Initialise and run a simple R project workflow},
  version = {1.0.0},
}
```
